---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Diana Korchmar (12362412)"
date: "`r Sys.Date()`"
output: distill::distill_article
---

## Exercise 1: Initializing git (4 Points)

a)  Start by initializing a new R project with git support, called `2024-exeRcise-sheet-3`.

**Status**: Done.

b)  Commit the files generated by Rstudio.

**Status**: Done.

c)  For all of the following tasks in this exercise sheet we ask you to always commit your changes after finishing each subtask e.g. create a commit after task *1d*, *1e* etc.

**Status**: Done.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a commit after this answer, see *1c*)

    **Strengths**

    • Very hard to lose files with git

    • History allows you to go back and understand change

    **Weaknesses**

    • Can be a bit complicated to use (esp. at first)

    • History takes up file space (only little)

e)  Knit this exercise sheet. Some new files will automatically be generated when knitting the sheet e.g. the HTML page. Ignore these files, as we only want to track the source files themselves.

**Status**: Done.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

a)  Create a new repository on GitHub in your account named `exeRcise-sheet-3`. Make sure you create a **public repository** so we are able to see it for grading. Add the link to the repository below:

    <https://github.com/dikorchmar/exeRcise-sheet-3>

b)  Push your code to this new repository by copying and executing the snippet on github listed under `…or push an existing repository from the command line`.

**Status**: Done.

c)  Regularly push your latest changes to GitHub again and especially do so when you are finished with this sheet.

**Status**: Done.

## Exercise 3: Baby-Names in Munich (3.5 Points)

a)  Download the data for both years and track it in git. For small datasets like these adding them to git is not a problem.

**Status**: Done.

b)  Load the data for both years into R. Check the type of the count variable ("Anzahl") and look into the data to determine why it is not numeric? Fix the problem in an appropriate manner, it is OK if some of the counts are inaccurate because of this. Explain your solution and the repercussions.

Loading the data for both years into R:

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
library(dplyr)
library(readr)

vornamen_2023 <- read_csv(file.path("vornamen-muenchen-2023.csv"))
vornamen_2022 <- read_csv(file.path("open_data_portal_2022.csv"))
```

Checking the type of the count variable ("Anzahl"):

```{r}
class(vornamen_2023$Anzahl)
class(vornamen_2022$Anzahl)
```

Converting "Anzahl" to numeric, removing any non-numeric characters:

```{r}
vornamen_2023$Anzahl <- as.numeric(gsub("[^0-9]", "", vornamen_2023$Anzahl))
vornamen_2022$Anzahl <- as.numeric(gsub("[^0-9]", "", vornamen_2022$Anzahl))
```

Checking if conversion was successful:

```{r}
class(vornamen_2023$Anzahl)
class(vornamen_2022$Anzahl)
```

Explaining my solution and the repercussions:

I cleaned the data by removing non-numeric characters from the "Anzahl" column using gsub("[\^0-9]", "", data\$Anzahl) and then converting these cleaned strings to numeric values with as.numeric(). This ensures all values are numeric, allowing for easier calculations, aggregations, and accurate sorting of the data.

However, this method can introduce slight inaccuracies. For example, entries like "4 oder weniger" are converted to 4, which may not reflect the exact count. While this simplifies the analysis and sorting, it's important to note that some counts might be slightly higher or lower than their actual values. Despite this, the overall trends and insights derived from the data remain reliable.

c)  Calculate the total number of babies born in Munich in 2022 and 2023. Which year had the bigger baby-boom?

```{r}
total_2022 <- sum(vornamen_2022$Anzahl, na.rm = TRUE)
total_2023 <- sum(vornamen_2023$Anzahl, na.rm = TRUE)

cat("Total babies born in 2022:", total_2022, "\n")
cat("Total babies born in 2023:", total_2023, "\n")

if (total_2022 > total_2023) {
  cat("2022 had the bigger baby-boom.\n")
} else {
  cat("2023 had the bigger baby-boom.\n")
}
```

d)  Add a new column `year` to both datasets which holds the correct year for each.

```{r}
vornamen_2022 <- vornamen_2022 %>% mutate(Year = 2022)
vornamen_2023 <- vornamen_2023 %>% mutate(Year = 2023)
```

e)  Combine both datasets into one using `bind_rows()`.

```{r}
combined_vornamen <- bind_rows(vornamen_2022, vornamen_2023)
```

f)  Combine the counts for same names to determine the most popular names across both years. Print out the top 10 names in a nicely formatted table for both years. Include a table caption.

There was no information on whether the column 'Geschlecht' needed to remain, so I also added it to the printed table with the top 10 names.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)

popular_names <- combined_vornamen %>%
  group_by(Vorname, Geschlecht) %>%
  summarise(total_count = sum(Anzahl)) %>%
  arrange(desc(total_count))

top_ten_names = head(popular_names, 10)

kable(top_ten_names, caption = "Top 10 most popular baby names in Munich across the years 2022 and 2023")

```

## Exercise 4: Open Analysis (4 points)

a)  Go to <https://github.com/owid/owid-datasets/tree/master/datasets> and choose a dataset that interests you. You can have a look at <https://ourworldindata.org/> to gather some inspiration.

b)  Download the dataset and track it in git.

c)  Put the name / title of the dataset and a link to it below.

-   Dataset Name: ...
-   Link: <https://github.com/owid/owid-datasets/>...

d)  Come up with a (research) question you want to answer with the data and briefly explain why you believe this is an interesting question within one sentence. It should be a question that can be answered with the dataset and using R.

e)  Use R to answer your chosen question.

f)  Create a meaningful plot / figure with the dataset. Make sure to provide a figure caption (via the chunk options / Rmarkdown) and correctly label the figure.
